---
title: "Data Manipulation: R package dplyr"
author: "Chih-Hui Wang (Jason)"
date: "April 28, 2015; Revised: March 10, 2016"
output: 
  html_document: 
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment='')
knitr::opts_chunk$set(fig.align='center')
```

This is an introduction to the `R` package `dplyr`, written by Hadley Wickham. Personally, I found it incredibly easy and useful to use `dplyr` when I did data manipulation. It's usually the first package I call when I start a data analysis project. The functions in the package work similar to SQL syntax. You almost can find any database function you want in `dplyr`. We will start by introducing the class `tbl`, the basic functions (verbs) to other special functions. We will also talk about how the amazing syntax `%>%` (like the pipe in Unix) makes your code more elegant.

```{r package, message=FALSE}
library(dplyr)
library(hflights)
library(rbenchmark)
```

We use `hflight` dataset for demo.
```{r data}
data(hflights)
names(hflights)
```

### 1. Data Type: tbl (table)

First of all, we introduce the class `tbl`. It will be better if we change the class to `tbl`. When you print the data.frame, `R` will dump everything to the console, even the data.frame is too big to fit in your console. However, if you convert the class to `tbl` and print it, it will give you a better output format. Note that the function in `dplyr` can be used to data.frame as well. We will compare the difference between `tbl` and data.frame later.
```{r tbl}
# hflights_df is data.frame
hflights_df <- hflights
class(hflights_df)

# Transform data.frame to table
hflights <- tbl_df(hflights)
class(hflights)

# Nicer output format
hflights
```

### 2. Verb 1: filter (Row Operation)

Subset is one of the most important things in data manipulation. In base `R`, we will use `subset` with conditions or just boolean vector with square brackets `[` to retrieve the rows we want. In `dplyr`, we use `filter(data, conditions)` to subset the data.

Let's say that we want to get the flights in March.
```{r filter}
# In dplyr
filter(hflights, Month == 3)

##### The following codes will give your the same result #####
# subset(hflights, Month == 3)
# hflights[hflights$Month == 3, ]
```
If you have multiple condition, you can use commas to separate several conditions. For example, `filter(hflights, Month == 3, DayofMonth == 16)` will pull out the filght on March 16. Also, you can use `&` and `|` operation in the arguments like `filter(hflights, Month == 3 & DayofMonth == 16)`.

We use the `benchmark` function in package `rbenchmark` to compare four methods. We replicate the procedure 100 times and calculate the time. You can also use the `system.time` to compare method while its result varies. `benchmark` gives us a more robust comparison.    
From the results, you can find that `filter` with class `tbl` is the fastest. As we mentioned earlier, the class `tbl` will be better than data.frame for large dataset. 
```{r benchmark}
# compare four methods
benchmark(
  method1=filter(hflights, Month == 3),
  method1_df=filter(hflights_df, Month == 3),
  method2=subset(hflights, Month == 3),
  method3=hflights[hflights$Month == 3, ],
  replications=100,
  columns=c("test", "elapsed", "relative", "user.self", "sys.self")
)
```
### 3. Verb 2: select (Column Operation)

We can use `select` function to select the column. We just put the column name in the arguments.
```{r select}
# In dplyr
select(hflights, Month, DayofMonth, FlightNum)
```

If you have a lot of column names shared some characters, without typing the explicit name you can use `contains`, `starts_with` and `ends_with` to pull out those columns. (Note that we do not show the result here. You can try the code.)
```{r select_function, eval=FALSE}
# contains
select(hflights, contains("delay"))

# starts_with
select(hflights, starts_with("M"))

# use ends_with
select(hflights, ends_with("th"))
```

### A nicer way to write codes in dplyr

Sometimes we will combine the `filter` and `select` function together. You may want to find out the flight number and Origin of those American Airlines (AA) flights. You will write code like the following:
```{r messy, eval=FALSE}
filter(select(hflights, FlightNum, Origin, UniqueCarrier), UniqueCarrier == "AA")
```
When you see the above codes, it may take a while to get what the codes is doing. However, in `dplyr`, it gives you a more elegant way to organize their code. It works like the pipe in Unix. We can rewrite the code as following: 
```{r pipe}
# Clearer way to write code
hflights %>%
    select(FlightNum, Origin, UniqueCarrier) %>%
    filter(UniqueCarrier == "AA")
```
You can interpret the `%>%` syntax as "then". The above code will be like that we first get the hflights data, select FlightNum, Origin, UniqueCarrier, and then filter the outcome to find those UniqueCarrier is AA. It throws the output in front of it to the argument of next function. In the following demo, I will use `%>%`.

### 4. Verb 3 arrange

Suppose we want to find out which airlines had the longest flights delayed in the data. We have to sort our data by the delay time. To do it in `dplyr`, we can use `arrange` function.
```{r arrange}
# To sort decreasingly, use desc
hflights %>%
  select(DepDelay, UniqueCarrier) %>%
  arrange(desc(DepDelay))
```

### 5. Verb 4 mutate

Sometimes, you may want to create a new variable. For instance, if we want to create the date variable in flight data, use `mutate` function.
```{r mutate}
# Create a new variable, date
hflights <- hflights %>% 
  mutate(date=paste(Year, Month, DayofMonth, sep="-"))

hflights %>% select(date)
```

### 6. Verb 5 summarise + Verb 6 group_by

Often, we will want to do something based on several groups such as what is the average delayed time for each airline. In base `R`, we use `tapply(hflights$DepDelay, hflights$UniqueCarrier, mean, na.rm=TRUE)` or `aggregate(DepDelay ~ UniqueCarrier, hflights, mean, na.rm=TRUE)`.
```{r group_by_summarise}
# Use group_by + summarise
hflights %>%
  group_by(UniqueCarrier) %>%
  summarise(average_delay=mean(DepDelay, na.rm=T))
```
Sometimes, we may want to apply the function to several columns or apply different functions to column.
```{r summarise_each}
#Apply function to multiple columns
hflights %>%
  group_by(UniqueCarrier) %>%
  summarise_each(funs(mean(., na.rm=T)), DepDelay, ArrDelay)

#Apply multiple functions to one column
hflights %>%
  group_by(UniqueCarrier) %>%
  summarise_each(funs(mean(., na.rm=T), min(., na.rm=T), max(., na.rm=T)), DepDelay)
```

### 7. Others

In this section, we are going to introduce some useful functions that can facilitate our analysis. The first one is ` n()`. It can help us count the number in the group. For example, we want to calculate the number of flights for each airline.  
```{r count_n}
#In dplyr
hflights %>%
  group_by(UniqueCarrier) %>%
  summarise(flight_count=n())
```

Suppose we also want to sort the airlines by the number of flights. We can add `arrange(fligh_count)` after `summarise`. Alternative is to use `count` function with argument `sort=TRUE` or, another useful function can count and sort at the same time:
```{r tally}
#Another easier way to write it
hflights %>%
  group_by(UniqueCarrier) %>%
  tally(sort=TRUE)
```

`n_distinct` funtion provides the number of unique subjects in a column.
```{r n_distince}
hflights %>%
  group_by(Dest) %>%
  summarise(flight_count=n(), plane_count=n_distinct(TailNum))
```

To check the data types of each column, use `glimpse`.
```{r glimpse}
#Like the str function in base R
glimpse(hflights)
```

We can also do sampling easily either by number or fraction.
```{r sample, eval=FALSE}
# Sample by number
hflights %>% sample_n(10)

# Or by fraction
hflights %>% sample_frac(0.25, replace=T)
```

# 8. Connecting with database

* dplyr can connect to a database as if the data was loaded into a data frame
* [Instruction for create a database](dplyr can connect to a database as if the data was loaded into a data frame)
```{r, eval=FALSE}
# connect to an SQLite database 
my_db <- src_sqlite("my_db.sqlite3")

# connect to the "hflights" table in that database
flights_tbl <- tbl(my_db, "hflights")

# identical query using the database
flights_tbl %>%
    select(UniqueCarrier, DepDelay) %>%
    arrange(desc(DepDelay))

# ask dplyr for the SQL commands
flights_tbl %>%
    select(UniqueCarrier, DepDelay) %>%
    arrange(desc(DepDelay)) %>%
    explain()
```

# References
* [Introduction to dplyr on CRAN](https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html)

* [Hands-on dplyr tutorial for faster data manipulation in R](http://www.r-bloggers.com/hands-on-dplyr-tutorial-for-faster-data-manipulation-in-r/)

* [A more comprehensive and advanced tutorial](http://rpubs.com/justmarkham/dplyr-tutorial)