---
title: "Principal Component Analysis"
author: "Jason"
date: "2015/07/03"
output:
  html_document:
    fig_height: 5
    keep_md: yes
  pdf_document:
    latex_engine: xelatex
---
```{r setup, echo=FALSE}
knitr::opts_chunk$set(comment="")
knitr::opts_chunk$set(fig.align='center')
```

```{r package}
#Package
library(ISLR); library(ggplot2); library(GGally)
```

```{r data}
arrest <- USArrests
summary(arrest)
```

```{r EDA}
#The function is built in GGally
ggpairs(arrest, title="Pair plot of all variable")
```

# Step of Calculating PCA   
1. Scale the data
```{r scale}
scale_arrest <- scale(arrest)
```
2. Get the covariance matrix
```{r covariance}
cov_arrest <- cov(scale_arrest)
cov_arrest
```
3. Calculate eigenvalues and eigenvectors
```{r eigen}
eigen_arrest <- eigen(cov_arrest)
eigen_arrest
```
4. Derive the new data (scores)

Final Data=Row Feature Vector X Row Data Adjust
```{r scores}
scores <- t(t(eigen_arrest$vectors) %*% t(scale_arrest))

head(scores)
```
\newpage

5. Choose the number of Component

a. Proportion of variance

Random vector 
\begin{displaymath}
\mathbf{X} = \begin{bmatrix}
- X_{1} -\\ 
- X_{2} -\\ 
\cdots \\ 
- X_{n} -\\
\end{bmatrix}
\end{displaymath}
$$Cov(a'X)=a'Cov(X)a$$
$\phi_{1}$ is the first eigenvector
$$Var(\phi_{1}'X)=\phi_{1}'Var\phi_{1}=\phi_{1}'\lambda_{1}\phi_{1}=\lambda_{1}\phi_{1}'\phi_{1}=\lambda_{1}$$
```{r show}
t(eigen_arrest$vectors[, 1]) %*% var(scale_arrest) %*% eigen_arrest$vectors[, 1]
```

```{r show2}
eigen_arrest$values[1]
```
After the project the original data to the component one. We can explained the origianl variation by eigenvalue1, $\lambda_{1}$. Hence, we can use the eigenvalue to see how much variation the component explain.
```{r eigen_way}
#Way1
eigen_value <- eigen_arrest$values
m <- data.frame(PC=1:4, PV=cumsum(eigen_value)/sum(eigen_value))

#plot
ggplot(m) + geom_line(aes(x=PC, y=PV), size=1, color="dodgerblue") +
  labs(title="Explained Proportion", x="# of Principal Component", y="Proportion of Variane") +
  theme_bw()
```

```{r score_variance}
#Way2
PV_each <- apply(scores^2, 2, sum)/sum(scale_arrest^2)
PV <- cumsum(PV_each)
m <- data.frame(PC=1:4, PV=PV)

#plot
ggplot(m) + geom_line(aes(x=PC, y=PV), size=1, color="dodgerblue") +
  labs(title="Explained Proportion", x="# of Principal Component", y="Proportion of Variane") +
  theme_bw()
```

b. Scree plot
```{r scree_plot}
eigen_value <- eigen_arrest$values
data <- data.frame(Principal_component=1:4, eigenvalue=eigen_value)
#plot
ggplot(data) + geom_line(aes(x=Principal_component, y=eigen_value), size=1, color="dodgerblue") +
  labs(title="Scree plot", x="Principal Component", y="Eigen vlaue") +
  theme_bw() + geom_hline(yintercept=1, linetype=2, color="red")
```

# Default function

There are two default functions to do Principal Component Analysis in R, which is `princomp` and `prcomp`. In the following discussion, we will split them into different section and compare their difference.   


# princomp
```{r PCA1}
pca1 <- princomp(scale_arrest)
```

```{r summary1}
summary(pca1)
```

```{r eigen_value}
#eigen value
pca1$sdev^2
```

```{r}
#loading
pca1$loadings
```

```{r}
#Score
head(summary(pca1)$score)
```

```{r biplot, fig.width=8, fig.height=7}
biplot(pca1)
```

# prcomp
```{r PCA2}
pca2 <- prcomp(scale_arrest)
```

```{r summary2}
summary(pca2)
```

```{r eigen_vlaue2}
pca2$sdev^2
```

```{r}
#loading
pca2$rotation
```

```{r}
#Score
head(scale_arrest %*% pca2$rotation)
```

Reference
* [Elementary Matrix Algebra Review](http://courses.washington.edu/b533/lect3.pdf)