---
title: "ggplot2"
author: "Jason"
date: "Monday, May 25, 2015"
output:
  html_document: 
    keep_md: yes
  pdf_document:
    fig_height: 4
    fig_width: 4
    latex_engine: xelatex
mainfont: Calibri
---
```{r setup, echo=FALSE}
knitr::opts_chunk$set(comment="")
```

http://zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/
For `ggplot`, `gg` is `Grammar of Graphics`. Think Verb, Noun, adjective for graphics. Plots are made up of *aesthetics* (size, shap, color) and *geoms* (points, lines) 

```{r}
library(ggplot2)

#Demo data
data(mtcars)
str(mtcars)
```

\newpage

#1. Basic function qplot
```{r, fig.align='center'}
#Fundamental
qplot(hp, mpg, data=mtcars)
```

```{r, fig.align='center', fig.width=5}
#Add color
#Type 1
qplot(hp, mpg, data=mtcars, color=am)
```

```{r, fig.align='center', fig.width=5}
#Type 2
qplot(hp, mpg, data=mtcars, color=as.factor(am))
```

```{r, fig.align='center', fig.width=5}
#Change shape
qplot(hp, mpg, data=mtcars, shape=as.factor(am))
```

```{r, fig.align='center'}
#Add geoms
qplot(hp, mpg, data=mtcars, geom=c("point", "smooth"), method="lm")
```

```{r, fig.align='center', fig.width=5}
#Histogram
qplot(mpg, data=mtcars, fill=as.factor(am), binwidth=2)
```

```{r, fig.align='center'}
#Facets
qplot(hp, mpg, data=mtcars, facets=. ~ am)
```

```{r, fig.align='center'}
#Facets
qplot(mpg, data=mtcars, facets=am ~ ., binwidth=2)
```

```{r, fig.align='center', fig.width=5}
#Density Smooth
qplot(mpg, data=mtcars, geom="density")
qplot(mpg, data=mtcars, geom="density", color=as.factor(am))
```

The qplot is quite similar to the plot in base R. In the following, we will introduce the ggplot function to make a more well-structrue and beautiful plot.
```{r, message=FALSE}
library(dplyr)

mtcars <- mtcars %>%
            mutate(cyl=as.factor(cyl)) %>%
            mutate(vs=as.factor(vs)) %>%
            mutate(am=as.factor(am))
str(mtcars)
```

# Scatter plot
```{r, fig.align='center'}
ggplot(mtcars) + geom_point(aes(x=hp, y=mpg))
```

```{r}
#Color by am
ggplot(mtcars) + geom_point(aes(x=hp, y=mpg, color=am))
```

```{r}
#Color by am and add trend line
ggplot(mtcars) + geom_point(aes(x=hp, y=mpg, color=am)) +
  geom_smooth(aes(x=hp, y=mpg, color=am), method="lm")
```

```{r}
#Divide into two facet
ggplot(mtcars) + geom_point(aes(x=hp, y=mpg, color=am)) +
  facet_grid(. ~ am) + geom_smooth(aes(x=hp, y=mpg, color=am), method="lm")
```

# Histogram
```{r}
ggplot(mtcars) + geom_histogram(aes(x=hp), binwidth=5)
```

```{r}
ggplot(mtcars) + geom_histogram(aes(x=hp, fill=am), binwidth=5, position="dodge")
```

```{r}
#Divide into two part
ggplot(mtcars) + geom_histogram(aes(x=hp, y=..density.., fill=am), binwidth=5, position="dodge") +
  facet_grid(am ~ .)
```
# Density plot
```{r}
ggplot(mtcars) + geom_density(aes(x=hp, fill=am)) +
  facet_grid(am ~ .)
```
# Density + histogram
```{r}
ggplot(mtcars) + geom_histogram(aes(x=hp, y=..density.., fill=am), binwidth=5) + 
  geom_density(aes(x=hp, color=am)) +
  facet_grid(am ~ .)
```

** Reference **

1. [ggplot themes](http://docs.ggplot2.org/0.9.2.1/theme.html)

2. [basic Intro to qplot](http://www.r-bloggers.com/basic-introduction-to-ggplot2/)

3. [Cheatsheet for Visualizing Distributions](http://www.r-bloggers.com/ggplot2-cheatsheet-for-visualizing-distributions/)