---
title: "Package: reshape2"
author: "Jason"
date: "2015年7月8日"
output: 
  html_document: 
    keep_md: yes
---
```{r setup, echo=FALSE}
knitr::opts_chunk$set(comment="")
knitr::opts_chunk$set(fig.align='center')
```

```{r package, message=FALSE}
library(reshape2); library(ggplot2); library(dplyr)
```

`reshape2` is a package that provide functions for us to manipulate with the data. It is very good at transforming the data from *wide* format to *long* format or otherwise. 

The **wide** format is shown below. It means that a column in the data represent one variable/feature of the data. 
```{r demo1}
data.frame(Var1=1:3, Var2=4:6, Var3=7:9)
```
And this is the **long** format
```{r demo2}
melt(data.frame(Var1=1:3, Var2=4:6, Var3=7:9))
```
Owing to the processing(dplyr) or grahp(ggplot), sometimes it will be very useful to transform the wide format data to long format. `reshape2` mainly provide two function to deal with two situation.

wide to long: `melt`
long to wide: `cast`
---

# melt
This is a toy example that I creat to use to illustrate the `melt`.
```{r toydata}
#Toy data
our_team <- data.frame(name=c("Jason", "Oracle", "Sanly"),
                height=c(182, 176, 160),
                weight=c(75, 77, 45))
our_team
```
Q1. Plot the variable `height` & `weight` together and use the color to distinguish two variable.
```{r melt}
m <- melt(our_team, id="name")
m
```
It mainly keep the variable you set in `id= ` and transform other variables to two column, one for varialbe and one for value.
```{r Graph}
ggplot(m) + geom_line(aes(x=name, y=value, group=variable, color=variable), size=1) +
  labs(title="CLW_R_Learning Group") +
  scale_color_manual(values=c("#1E90FF", "#FF8C00"))
```

#cast
```{r toydata2}
our_team2 <- data.frame(name=rep(c("Jason", "Oracle", "Sanly"), 4),
                year=rep(c(2012, 2013, 2014, 2015), each=3),
                height=c(180:183, 173:176, 157:160),
                weight=c(85, 80, 75, 75, 60, 70, 75, 77, 47, 46, 45, 45))
our_team2
```

```{r melt2}
m2 <- melt(our_team2, id=c("name", "year"))
m2
```
You can transform the data back to original by dcast.
```{r dcast}
#Make it back to origin
d <- dcast(m2, name + year ~ variable)
d
```
Another way to do it.
```{r another}
dcast(m2, ... ~ variable)
```

It also includes some aggregation funciton.
```{r other_way}
dcast(m2, name ~ variable, mean)
```
Use `margins` to calculate all value for each row and column
```{r}
dcast(m2, name ~ variable, mean, margins=TRUE)
```
If you just want the margin of name:
```{r}
dcast(m2, name ~ variable, mean, margins="variable")
```
You can get the same answer by using dplyr as below.
```{r dplyr}
our_team2 %>%
  group_by(name) %>%
  summarise(height=mean(height), weight=mean(weight))
```
You can also use the `acast` to obtain a higher dimensional array. (Change the order in formula will give you different result, you can play it around)
```{r}
acast(m2, name ~ year ~ variable)
```

